---
title: "rsv_model"
author: "David Hodgson"
date: "24/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Setup 
# 1.1 Import libraries
May need to downoad all of these (use install.packages)
```{r libraries}
library(Rcpp)       # For c++ integration
library(RcppEigen)  # Ditto
library(coda)       # Ditto
library(tidyverse)
library(parallel)
library(MASS)
library(foreach)
library(gridExtra)
library(doParallel)
library(BayesianTools)

file.path = 'rsv_model_R-master' # change as appropriate
source(paste0(file.path, '/src/helpers.R')) # some important helper functions
```


## 2 Import data for the epidemic model 
Upload the data used to fit he model, include the contact matrices, the fitted parameter value's upper and lower limits, the observational data (which is NOT the same as in the paper).
```{r import data}

load(file=paste0(file.path, '/Rdata/rsv_data.Rda'))   # Load in model data, including observation data (which is made up (not the same as I used in the paper).

```

## 3. Import that model from the cpp file and update the parameter values 
Once the data is loaded import the model from cpp and change the parameter values.
Need to link install boost (easiest to do via homebrew), might throw loads of [-Wunknown-pragmas] warnings, just ignore, usually an issue with the coompiler.
```{r}

sourceCpp(paste0(file.path, "/src/logLikelihoodModule.cpp")) # ensure c++14 is enabled
classEvaluateLogLikelihood <- new(EvaluateLogLikelihood, numberDailyLiveBirths, population, ageGroupBoundary) # Calls class
classEvaluateLogLikelihood$contactMatrixPhy <- contactMatrixPhy 
classEvaluateLogLikelihood$contactMatrixCon <- contactMatrixCon
classEvaluateLogLikelihood$observedData <- as.matrix(observationalData)
classEvaluateLogLikelihood$lowerParamSupport <- fit.par$lowerParSupport
classEvaluateLogLikelihood$upperParamSupport <- fit.par$upperParSupport
classEvaluateLogLikelihood$run_full <- nrow(observationalData)*7 # number of days to fit the data and model to

```

## using BayesianTools package

```{r}

# Get the log likelihood
llikelihood = function(params){
  ll <- classEvaluateLogLikelihood$evaluateLogLikelihoodCpp(params)
  if (ll == -Inf) # This is to stop infinite log likelihoods gettings stuck
    ll = -1000000
  return(ll)
}

# Get the log priors (density and samplers)
density = function(params){
  p=0
  for (i in 1:length(fit.par$namesOfParameters)){
      if (params[i] < fit.par$lowerParSupport[i] || params[i] > fit.par$upperParSupport[i]){ return(log(0)) }
      else{
          p <- p + prior(params[i], fit.par$lowerParSupport[i], fit.par$upperParSupport[i], priorDistDensity[[i]], TRUE, priorDistPar1[i], priorDistPar2[i])
      }
  }
  return(p)
}

sampler = function(n=1){
    s <- vector()
    for (i in 1:length(fit.par$namesOfParameters)){
      s[i] <- prior.sample(fit.par$lowerParSupport[i], fit.par$upperParSupport[i], priorDistSample[[i]], priorDistPar1[i], priorDistPar2[i])
    }
    s
}

priors <- createPrior(density=density, sampler=sampler, lower=fit.par$lowerParSupport, upper = fit.par$upperParSupport)

# Create Bayesian Setup
bayesianSetup <- createBayesianSetup(llikelihood, priors, names = rownames(fit.par$namesOfParameters))
```

```{r}

# Differential Evoluation with Snooker
settingsDezs <- list(iterations = 1000000, nrChains = 3, thin = 10)
outDezs <- runMCMC(bayesianSetup = bayesianSetup, sampler = "DEzs", settings = settingsDezs) 

```